Tulai

### 2023-09-08 Tulai.qmd
test

# Check unique values in the modified columns
print(unique(tl4$SHINE))

@@

# Then replace 'NA', 'na', or '0' with R's NA
tl4 <- tl4 %>%
  mutate(across(everything(), ~na_if(., "NA"))) %>%
  mutate(across(everything(), ~na_if(., "na"))) %>%
  mutate(across(everything(), ~na_if(., "0")))

# Show the updated data frame
print(tl4)

@@

@@

```{r}
# Create a summary table based on the 'al' column
summary_table_al <- tl_final %>%
  group_by(al) %>%  
  summarise(
    `Cores (n)` = sum(!is.na(core)),  # Count non-NA core entries
    `Core rejuvenations pieces (n)` = sum(!is.na(corefrag)),  # Count non-NA corefrag entries
    `Blade (n)` = sum(blank == "blade", na.rm = TRUE),
    `Flake (n)` = sum(blank == "flake", na.rm = TRUE),
    `Bladelet (n)` = sum(blank == "bladelet", na.rm = TRUE),
    `Retouched Tools (n)` = sum(retouch == 1, na.rm = TRUE),
    `Utilized (n)` = sum(utilization == 1, na.rm = TRUE),
    `Sickle Shine (n)` = sum(`sickle shine` == 1, na.rm = TRUE)
  )

@@

```{r}
# Create a tibble with example data
summary_table2_area <- tibble(
  area = c("D1_a", "D1_c", "TA_d", "TA_e", "TP1_c"),
  lithic_volumetric_density = c(51.2, 110.9, 45.1, 100.2, 88.3),
  percent_retouch = c(4.9, 9.3, 15.3, 18.0, 9.8)
)

@@

# Create the ggplot
p <- ggplot(summary_table2_area, 
            aes(x = lithic_volumetric_density,
                                     y = retouch_frequency)) +

@@

print(p)

@@

data_sheet <- "https://docs.google.com/spreadsheets/d/16yLb-RyIQ7x2jFEH_Q7OZYTwlEAXJgFu/edit#gid=2014421097"

sum_depth_sheet <- read_sheet(data_sheet, sheet = "sum_depth")

@@

# Plotting
p <- ggplot(summary_table2_area, aes(x = lithic_volumetric_density, y = percent_retouch)) +

@@

print(p)

@@

library(dplyr)
library(ggplot2)

# Filter the data for the relevant categories
tl4_filtered <- tl4_tool_Blank %>%
  filter(Blank2 %in% 
           c("bladelet", 
             "blade", 
             "flake"))

# Create separate bar plots for shine, Retouch, and Utilization
plot_shine2 <- ggplot(tl4_filtered, aes(x = Blank2, fill = shine2)) +
  geom_bar(position = "dodge") +
  labs(title = "Shine2 by Blank2",
       x = "Blank2",
       y = "Frequency",
       fill = "Shine2") +
  theme_minimal()

plot_retouch <- ggplot(tl4_filtered, aes(x = Blank2, fill = Retouch)) +
  geom_bar(position = "dodge") +
  labs(title = "Retouch by Blank2",
       x = "Blank2",
       y = "Frequency",
       fill = "Retouch") +
  theme_minimal()

plot_utilization <- ggplot(tl4_filtered, aes(x = Blank2, fill = Utilization)) +
  geom_bar(position = "dodge") +
  labs(title = "Utilization by Blank2",
       x = "Blank2",
       y = "Frequency",
       fill = "Utilization") +
  theme_minimal()

# Combine the three plots side by side
library(patchwork)
combined_plots <- plot_shine2 + plot_retouch + plot_utilization
combined_plots

```{r}
library(dplyr)
library(knitr)
library(kableExtra)

#table

tl_final_aread <- tl_final_area %>%
  filter(blank %in% c("bladelet", "blade", "flake"))  %>%

  group_by(blank, utilization, `sickle shine`, retouch) %>%
  summarise(count = n())

table_kable <- kable(table_summary, format = "html", caption = "Summary Table", 
                     col.names = c("Blank2", "Utilization", "shine2", "Retouch", "Count")) %>%
  kable_styling(bootstrap_options = "striped")

print(table_kable)

```{r}
#|label: fig_
#|fig-cap: "Boxplot for each Dimention"
#|fig-height: 5

library(dplyr)
library(ggplot2)

# Filter the data to keep only bladelet, blade, and flake in the Blank column
tl4_tool_Blank_filtered <- tl4_tool_Blank %>%
  filter(Blank2 %in% c("bladelet", "blade", "flake"))

tl4_tool_Blank_tally <- 
  tl4_tool_Blank_filtered %>%
  group_by(Blank2) %>%
  tally()

# Select dimensions and convert Thickness (mm) to numeric
tl4_tool_Blank_selected <- 
  tl4_tool_Blank_filtered %>% 
  select(Blank2, 
         Utilization, 
         Retouch, 
         shine2)

tl4_tool_Blank_long <- 
  tl4_tool_Blank_selected %>% 
  pivot_longer(cols = c(Utilization, Retouch, shine2), 
               names_to = "Blank2", 
               values_to = "Value")

# Create boxplot
ggplot(tl4_tool_Blank_long, aes(x = Blank2, y = Value)) +
  geom_boxplot() +
  facet_wrap(~Blank2, scales = "free") +
  labs(x = "Blank", 
       y = "Frequency") +
  theme_minimal()

```{r}
library(dplyr)
library(ggplot2)

# Filter the data to keep only bladelet, blade, and flake in the Blank2 column
tl4_tool_Blank_filtered <- tl4_tool_Blank %>%
  filter(Blank2 %in% c("bladelet", "blade", "flake"))

tl4_tool_Blank_tally <- 
  tl4_tool_Blank_filtered %>%
  group_by(Blank2) %>%
  tally()

# Select dimensions and convert Thickness (mm) to numeric
tl4_tool_Blank_selected <- 
  tl4_tool_Blank_filtered %>% 
  select(Blank2, 
         Utilization, 
         Retouch, 
         shine2)

# Manually rename the duplicate columns
names(tl4_tool_Blank_selected)[names(tl4_tool_Blank_selected) == "Blank2"] <- "Dimension"

# Pivot data to longer format
tl4_tool_Blank_long <- 
  tl4_tool_Blank_selected %>% 
  pivot_longer(cols = c(Utilization, Retouch, shine2), 
               names_to = "", 
               values_to = "Value")

ggplot(tl4_tool_Blank_long, aes(x = Dimension, y = Value)) +
  geom_boxplot() +
  facet_wrap(~, scales = "free") +
  labs(x = "Blank", 
       y = "Frequency") +
  theme_minimal()

@@

# Load required packages
library(ggplot2)

# Create the plot with facetting
tl4_plots <- ggplot(tl4_filtered, aes(x = Blank2, fill = Blank2)) +
  geom_bar(position = "dodge") +
  labs(title = "Utilization, Retouch, and shine2 by Blank",
       x = "Blank2",
       y = "Count") +
  facet_grid(. ~ Blank2, scales = "free") +
  theme_minimal()

print(tl4_plots)

>>>>

print(unique(tl4$SHINE))

@@

# replace 'NA', 'na', or '0' with R's NA
tl4 <- tl4 %>%
  mutate(across(everything(), ~na_if(., "NA"))) %>%
  mutate(across(everything(), ~na_if(., "na"))) %>%
  mutate(across(everything(), ~na_if(., "0")))

print(tl4)

@@

```{r}
#set the value of utilization to NA only for the rows where `sickle shine` == 1. 
#All other rows should remain unchanged

#unsuccessful, should try again later

@@

```{r}
# Create a summary table based on the 'al' column
summary_table_al <- tl_final %>%
  group_by(al) %>%  
  summarise(
    `Cores (n)` = sum(!is.na(core)),  
    `Core rejuvenations pieces (n)` = sum(!is.na(corefrag)),  
    `Blade (n)` = sum(blank == "blade", na.rm = TRUE),
    `Flake (n)` = sum(blank == "flake", na.rm = TRUE),
    `Bladelet (n)` = sum(blank == "bladelet", na.rm = TRUE),
    `Retouched Tools (n)` = sum(retouch == 1, na.rm = TRUE),
    `Utilized (n)` = sum(utilization == 1, na.rm = TRUE),
    `Sickle Shine (n)` = sum(`sickle shine` == 1, na.rm = TRUE)
  )

@@

```{r}
# Create a tibble 
summary_table2_area <- tibble(
  area = c("D1_a", "D1_c", "TA_d", "TA_e", "TP1_c"),
  lithic_volumetric_density = c(51.2, 110.9, 45.1, 100.2, 88.3),
  percent_retouch = c(4.9, 9.3, 15.3, 18.0, 9.8)
)

@@

# Create the ggplot
WABI_AL <- ggplot(summary_table2_area, 
            aes(x = lithic_volumetric_density,
                                     y = retouch_frequency)) +

@@

print(WABI_AL)

@@

sum_depth_sheet <- "https://docs.google.com/spreadsheets/d/1Q0QZESk412ZQLE24yPs6Rg-7Y9OMX63DHoXeIPNAIFM/edit#gid=0"

sum_depth_sheet <- read_sheet(sum_depth_sheet)

sum_depth_sheet <- sum_depth_sheet %>%
  mutate(area_layer_level = paste0(toupper(area), "_", layer, "_", level)) %>%
  relocate(area_layer_level, .after = level)

sum_depth_sheet <- sum_depth_sheet %>%
  mutate(retouch_freq = round(retouch_tool / total_lithics, 2),
         lithic_dens = round(total_lithics / volume, 2)
  )

write_sheet(sum_depth_sheet, ss = sheet_url)

```{r}
# Create the ggplot
WABI_Level <- ggplot(sum_depth_sheet, 
            aes(x = lithic_dens,
                y = retouch_freq)) +

  geom_point(color = "black", size = 2) +

  geom_smooth(method = "lm", 
              se = TRUE, 
              color = "blue", 
              size = 0.5) +

  geom_text(aes(label = area_layer_level), 
            vjust = 1.5, 
            hjust = 0.5, 
            size = 2,
            check_overlap = TRUE) +

  stat_poly_eq(aes(label = paste(stat(eq.label), 
                                 stat(rr.label), 
                                 stat(p.value.label), 
                                 sep = "~~~")),
               formula = y ~ x, 
               parse = TRUE, 
               size = 2) +

print(WABI_Level)

```{r}
# Create the ggplot with y-axis on a log scale
WABI_Level <- ggplot(sum_depth_sheet, 
            aes(x = lithic_dens,
                y = retouch_freq)) +

  # Add points with different shapes based on area
  geom_point(aes(shape = area), size = 2) +

  # Add smoothing line
  geom_smooth(method = "lm", 
              se = TRUE, 
              color = "blue", 
              size = 0.5) +

  # Add regression equation, R-squared, and p-value
  stat_poly_eq(aes(label = paste(stat(eq.label), 
                                 stat(rr.label), 
                                 stat(p.value.label), 
                                 sep = "~~~")),
               formula = y ~ x, 
               parse = TRUE, 
               size = 2) +

  # Apply log scale to y-axis
  scale_y_log10() +

  # Add shape legend
  scale_shape_manual(values = c(16, 17, 25)) +

  # Customize theme
  theme(
    axis.text = element_text(size = 8),
    axis.title = element_text(size = 10),
    plot.title = element_text(size = 12)
  )

# Print the plot
print(WABI_Level)

```{r}
# log-scale

WABI_Level <- ggplot(sum_depth_sheet, 
            aes(x = lithic_dens,
                y = retouch_freq)) +

  # Add smoothing line
  geom_smooth(method = "lm", 
              se = TRUE, 
              color = "blue", 
              size = 0.5) +

  # Add text labels for each point
  geom_text(aes(label = area_layer_level), 
            vjust = 1.5, 
            hjust = 0.5, 
            size = 2,
            check_overlap = TRUE) +

  stat_poly_eq(aes(label = paste(stat(eq.label), 
                                 stat(rr.label), 
                                 stat(p.value.label), 
                                 sep = "~~~")),
               formula = y ~ x, 
               parse = TRUE, 
               size = 2) +

  labs(
    title = "Relationship between Retouch Frequency and Lithic Volumetric Density",
    x = "Lithic Volumetric Density",
    y = "Retouch Frequency"
  ) +

  scale_y_log10() +

print(WABI_Level)

# Create the ggplot
WABI_Level <- ggplot(sum_depth_sheet, 
            aes(x = lithic_dens,
                y = retouch_freq)) +

  # line with log scale

stat_smooth(method = "lm", 
            formula = y ~ log(x), 
            se = TRUE, aes(group = 1), 
            color = "blue", 
            size = 0.5)

  # Add text labels for each point
  geom_text(aes(label = area_layer_level), 
            vjust = -1, 
            hjust = -0.5, 
            size = 2.5) +

print(WABI_Level)

@@

# Plotting
p_area <- ggplot(summary_table2_area, aes(x = lithic_volumetric_density, y = percent_retouch)) +

@@

print(p_area)

@@

@@
